---
title: "Exercise 2"
author: '[Your first name] [Your last name]'
date: "`r Sys.Date()`"
output: html_document
---

## Setup

Install the packages you'll need for this assignment if you haven't already.

```{r setup}
# Install pacman
if (!require("pacman")) install.packages("pacman")
# p_load function loads packages if installed, or install then loads otherwise
pacman::p_load(dplyr, readr)
```

## Question 1

### Data collection

#### About the data

For this question, you'll use 1-year total population estimates for King County, WA, from the American Community Survey (ACS). Think of these as estimates of total person-years lived during a particular calendar year. During this course, we'll assume for simplicity that there is no margin of error in these estimates, which is factually inaccurate.

Learn about ACS here: https://www.census.gov/programs-surveys/acs/about.html

#### Step 1: Download the data

1. Download the data from here: ENTER URL
2. Move the downloaded file into your R Project folder for this assignment

#### Step 2: Read the data into R

1. Since your working directory is your R Project folder, you can replace the `file_dir` string below to be an empty string `""`

```{r q1-read-data}
file_dir <- ".../"
file_name <- "king_acs1_totpop.csv"
file_path <- paste0(file_dir, file_name)
readr::read_csv(file_path)
```

